### 2020-08-28 latex/vlp.tex
Wrote the energy dissipation and stabilization results
Finished the bulk of th eVLP section, need to just add figures now and
then start on simulation.

We will show that anti-Lyapunov function exists for our VNHC.
In order to do so however, we need to prove the following lemma.

@@

\begin{thm}\label{thm:vlp-gains-energy}
   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pi,0)\right\}\).
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underbar{l},\bar{l}]\) such that 
   \[
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \]
   with the property that the inequality is strict for almost every \(\theta\).
\end{thm}
\begin{proof}
    Choose, as a candidate anti-Lyapunov function, the energy for the average-length pendulum 
    \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
                    + m g l_\text{avg} (1-\cos(q))
    \]
    which is positive definite and has derivative 
    \[
      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
                 {l_\text{avg}^2l(\theta)^2}
    \]
    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
    by Lemma \ref{lemma:sign-of-cube},
    \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
    \]
    Then the derivative of \(E_\text{avg}\) is almost always positive, since
    \begin{align*}
       \label{eq:}
       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                   &\geq 0 \text{ (by assumption)}
    \end{align*}
    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
    Since the initial condition is not an equilibrium,
    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
    origin and the VLP is gaining energy on average.
\end{proof}

\begin{cor}
   The VNHC \(h(q,p) = l - l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
   over time, since 
   \(\left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
   -\Delta l \sin(2\theta) < 0\) for almost every \(\theta\).
\end{cor}

% TODO: Describe energy reduction, which comes from switching the sign

Theorem \ref{thm:vlp-gains-energy} provides an alternate explanation for why the
optimal controller \(l^\star(\theta)\) works so well at injecting energy: it
maximizes the derivative of \(E_\text{avg}\) under the restriction 
\(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
\((q,p)\)-plane diverges from the origin as fast as possible. The differentiable
controllers \(l_T(\theta)\) will also inject energy, albeit at a rate which
decreases as \(T\) increases. That said, Theorem \ref{thm:vlp-gains-energy} and the
subsequent Corollary shows that VNHCs are a powerful tool in creating effective energy
pumping strategies based on natural human motion.

% TODO: How would we stabilize a specific energy level?

>>>>

We will show that anti-Lyapunov function exists for our VNHC.
In order to do so, we require the following lemma.

@@

\begin{thm}\label{thm:vlp-energy-stabilization}
   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pm\pi,0)\right\}\).
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underbar{l},\bar{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \end{equation}
   with the property that the inequality is strict for almost every \(\theta\).
   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) leads to
   energy dissipation.
\end{thm}
\begin{proof}
    Choose, as a candidate anti-Lyapunov function, the energy for the average-length pendulum 
    \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
                    + m g l_\text{avg} (1-\cos(q))
    \]
    which is positive definite at \((0,0)\) and has derivative 
    \[
      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
                 {l_\text{avg}^2l(\theta)^2}
    \]
    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
    by Lemma \ref{lemma:sign-of-cube},
    \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
    \]
    Then the derivative of \(E_\text{avg}\) is almost always positive, since
    \begin{align*}
       \label{eq:}
       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                   &\geq 0 \text{ (by assumption)}
    \end{align*}
    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
    Since the initial condition is not an equilibrium,
    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
    origin and the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
    means \(E_\text{avg}\) is a regular Lyapunov function, since its derivative
    will be negative almost everywhere.
\end{proof}

\begin{cor}
   The VNHC \(l = l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
   over time, since 
   \[
      \left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
   -\Delta l \sin^2(2\theta) \geq 0
   \] 
   Likewise, let 
   \(l\inv_\frac{\pi}{2}(\theta) := l_\text{avg} - \Delta l \sin(2\theta)\).
    The VNHC \(h(q,p) = l - l\inv_\frac{\pi}{2}(\theta)\) which flips 
    \(l_\frac{\pi}{2}(\theta)\) about \(l_\text{avg}\) will dissipate energy from
   the VLP, since 
   \[
      \left(l\inv_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
      -\Delta l \sin^2(2\theta) \leq 0
   \] 
\end{cor}

The types of VNHCs which satisfy Theorem \ref{thm:vlp-energy-stabilization} are illustrated by Figure
\ref{fig:vlp-energy-in-out}. To stabilize specific energy level sets, one simple
approach is to switch between injection and dissipation VNHCs depending on the
current energy level.
For the VNHCs we designed in this section, this means toggling
between \(l_\frac{\pi}{2}(\theta)\) and \(l\inv_\frac{\pi}{2}(\theta)\),
possibly with some hysteresis to avoid infinite switching.

\begin{figure}
   \centering
   % TODO: Add a figure showing regions for injection vs dissipation
   \caption{}\label{fig:vlp-energy-in-out}
\end{figure}

Theorem \ref{thm:vlp-energy-stabilization} also provides an alternate
explanation for why the ``optimal" VNHC \(l^\star(\theta)\) works so well at
injecting energy: it maximizes the derivative of \(E_\text{avg}\) under the
restriction \(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
\((q,p)\)-plane diverges from the origin as fast as possible.

Let us define \((l^\star)\inv(\theta)\) by swapping the order of 
\(\underbar{l}\) and \(\bar{l}\) in \(l^\star(\theta)\). Since this 
\textit{minimizes} the derivative of \(E_\text{avg}\) under the restriction
\(l(\theta) \in [\underbar{l},\bar{l}]\), we can predict that 
\((l^\star)\inv(\theta)\) is the VNHC representation of an optimal 
energy dissipation controller.
This is, in fact, true: \cite{pumping_swing_standing_squatting} showed 
that squatting at the lowest point of a swing and standing at the highest
point (instead of standing and squatting resp.) produces the
time-optimal trajectory for stopping a standing swing.

All together, these results show that VNHCs can replicate the time-optimal
strategy performed by humans in a time-independent manner. Furthermore, we see
that VNHCs are a powerful tool for creating simple energy stabilization
techniques based on natural human motion.
